Automatic Hadith Graphs using Arabic Morphological Analysis

### 2011-06-22 trunk/tex/hadith.ijcnlp.tex
some refactoring and correcting a bug in the rasoul words
also committing the tex file which had not been previously been so

\begin{abstract}
\end{list}
A hadith in Islamic literature is a narration from the 
prophet Mohammad related by multiple narrators.
Establishing the authenticity of a hadith is an important task
based on which Islamic scholars issue rules that affect the life
of people around the world. 
The literature lacks an automated and exhaustive mechanism of 
authentication checks. 
This is currently manual work prone to error and almost
impossible to complete in the life time of a scholar due to the
large literature available. 
In this paper, we present a tool that uses 
an Arabic application-specific morphological analyzer 
to successfully automate the
analysis of three books of hadith selected 
arbitrarily and abstract 
each book into a graph of narrators related to each other. 
\end{abstract}

@@

In this paper, we present the hadith {narrator graph
extractor using morphological analysis} (NaGEMA), 
a novel technique that automates
extracting the \noArRL{sanad} from the books of traditions 
into a diagram of chains of narrators. NaGEMA segments
a book into narrations, each narration into its \noArRL{matn} 
and its \noArRL{sanad}, 
and each sanad into its separate narrators. 
This work will lead to 
automate an exhaustive \noArRL{.hady_t} authentication 
effort in the future as we plan to 
automatically analyze the biography books.

@@

{\bf Sarf.~~~}
In this paper, 
we use Sarf~\footnote{Sarf (name modified for blind review) }
an {\em application-specific efficient
morphological analyzer} that uses 
non-deterministic finite state Automata 
driven by an application-specific controller customized 
by the user. 
%Each machine in Sarf takes one letter at a time as input
%and represents a valid morphological analysis.
The application-specific controller prunes false positives
early in the run by making a decision at every input letter
to control the machines.
Sarf builds on and extends the lexicon of 
Buckwalter~\shortcite{Buckwalter:02} 
with proper and location names extracted from different online 
sources~\footnote{\href{http://alasmaa.net/}{http://alasmaa.net/ }, 
\href{http://ar.wikipedia.org/}{http://ar.wikipedia.org/}}
%as well as biblical sources~\footnote{Genesis 4:17-23; 5:1-32; 9:28-10:32; 11:10-32; 25:1-4, 12-18; 36:1-37:2; Exodus 6:14-25; Ruth 4:18-22; 1 Samuel 14:49-51; 1 Chronicles 1:1-9:44; 14:3-7; 24:1; 25:1-27:22; Nehemiah 12:8-26; Matthew 1:1-16; Luke 3:23-38}.
as well as biblical sources.
%~\footnote{Genesis 4:17-23; 5:1-32; 9:28-10:32; 11:10-32; 25:1-4, 12-18; 36:1-37:2; Exodus 6:14-25; Ruth 4:18-22; 1 Samuel 14:49-51; 1 Chronicles 1:1-9:44; 14:3-7; 24:1; 25:1-27:22; Nehemiah 12:8-26; Matthew 1:1-16; Luke 3:23-38}.

@@

@@

%Several researchers~\cite{Hadithopaedia:08} attempted to automate 
%the analysis of the hadith literature. 
In this paper, we use Sarf to automate the
analysis of three books of hadith selected 
arbitrarily~\cite{IbnHanbal,AlKulayni,AlTousi}\footnote{We obtained
  the digitized books from online sources such as 
  \href{http://www.yasoob.com/}{http://www.yasoob.com/} and 
  \href{http://www.al-eman.com/}{http://www.al-eman.com/}. }.
%  Our understanding is that those books are historical documents with 
 % public IP.}.
%\setcode{utf8}
%\begin{arabtext}
%حدثنا \emphasize{هشيم}، أخبرنا \emphasize{إسماعيل بن أبي خالد}،
% عن أبي إسحاق، عن سعيد بن جبير، قال كنا
%    مع ابن عمر رضي الله عنه حيث أفاض من
%    عرفات إلى جمع فصلى بنا المغرب ومضى
%\end{arabtext}
%\setcode{standard}
NaGEMA accepts a book as input
and segments it into a vector of hadiths. 
NaGEMA segments each hadith into its
sanad and matn parts. 
NaGEMA 
detects the chain of narrators in the sanad and 
the relation that links each narrator to his ancestor and 
predecessor in the chain. 
It also detects the full name of each narrator that is
composed of several proper names with connectors
in between.

To do that NaGEMA queries Sarf for the following morphological 
categories.
\begin{itemize}
\item $\mathit{NAME}$: a proper name of a person such as \RL{'a.hmad} .
\item $\mathit{IBN}$: a special name connector meaning the son of such as \RL{ibn}.
\item $\mathit{NMC}$: other name connectors. 
\item $\mathit{NISBA}$: a possessive adjective that qualifies a person such 
as \RL{alma.sriy} ``the Egyptian''. 
\item $\mathit{NRC}$: a narrator connector such as
\RL{`an} ``on behalf of'', \RL{.hada_t} ``narrated'', \RL{qaal} ``said'', 
\RL{'a_hbar} ``told'' or one of their morphological equivalents. 
\end{itemize}

@@

@@

The process also benefits from name connector 
words and narrator connector words to enhance the
accuracy of the chain extraction and to segment the 
chain accurately into narrators.
Once we meet a name connector or a narrator 
connector word we reset the word counters that we
compare against the thresholds. 
that directly interacted with Sarf.

@@

We implemented our hadith chain extraction 
in an finite state machine
controller that interfaces with Sarf.
Intuitively, the controller reports a valid chain 
of narrators when a sequence of names
connected by narrator connectors appears. 
%This includes the task of finding multi-word names
%such as \RL{`bd alr.hmn} often appearing as ``run-on'' words.
The controller marks the beginning of the hadith with 
the beginning of 
the current proper name sequence,
and marks the end of the hadith with the beginning of the 
next proper name sequence. 
It detects the chain of narrators as the sequence itself.

@@

The controller moves to state $\mathit{NAME\_S}$ on
$\mathit{NAME}$.
It moves to state $\mathit{NRC\_S}$ 
when Sarf reports an $\mathit{NRC}$.
State $\mathit{NMC\_S}$
indicates that the controller expects a name to appear within 
a tolerance threshold expressed by 
$\theta_{\mathit{nmc}}$ and $\lambda_{\mathit{bmc}}$.
It returns to $\mathit{NAME\_S}$ when a name is met, 
loops when still within the threshold, and 
resets to $\mathit{TEXT\_S}$ when the threshold is exceeded 
signaling that the last collection of names met do not qualify
as a narrator or a chain or narrators.

@@

\section{The narrator graph}

@@

The directed acyclic graph (DAG) 
in Figure~\ref{f:narrators} shows a partial order relation (POR) 
between narrators.
We automatically extracted the POR from the three books of 
hadith~\cite{IbnHanbal,AlKulayni,AlTousi}.
The nodes in boxes are the \RL{matn} of the hadith, 
and the other nodes are the narrators.
Given a set of chains of narrators similar to 
Figure~\ref{f:exhadith} we formed the DAG using graph algorithms 
that merged equal names in the chains. 
We compared the narrators together using a morphological
distance function. 
We considered each narrator to have several morphological 
components and we rewarded every morphological match with 
a small $\delta$. 
Then we applied a threshold $\theta$ to the sum of rewards
and decided the equality of the narrators. 
%We omitted the details of the distance function for brevity.

\begin{table}[tb]
%\centering
%\resizebox{.9\columnwidth}{!}{
\begin{tabular} {p{6cm}}
\begin{Verbatim}[fontsize=\relsize{-2},
frame=topline,framesep=4mm,label=\fbox{MergeChains algorithm},
commandchars=\\\{\}, codes={\catcode`$=3\catcode`_=8}]
Merge(Chain $C$[])
  Chain $C_c$[]=\{\}
  foreach Chain $c_1$ in $C$ 
    $C_c = C_c \cup c_1$
    foreach Chain $c_2$ in $C-C_c$
      foreach Narrator $n_1$ in $c_1$
        start = max( $n_1$.index - radius, 0)
        end = min($n_1$.index + radius, $n_2$.size)
        $n_2$ = $c_2$[start]
        while (true)
            if (Distance($n_1$, $n_2$) < $\theta$)
                mergeNodes($n_1$, $n_2$)
            if ($n_2$ = $c_2$[end])
                    break
            $n_2$ = $c_2$.next($n_2$)        
\end{Verbatim}
\end{tabular}
%}
\label{a:merge}
\end{table}

@@

@@

  $I_1$ = extractIbn( stems(n1.connectors) )
  $I_2$ = extractIbn( stems(n2.connectors) )

  $C_1$ = stems(n1.connectors - $I_1$)
  $C_2$ = stems(n2.connectors - $I_2$)

@@

  if ($C_1 \cap$ Places $\not= C_2 \cap$ Places)
    dist += $\delta_3$
  else
    dist -= $\delta_3$

@@

  if (correctIbnCorrespondence($I_1$, $I_2$))
      dist -= $\delta_5$
  return (MAX - dist)
\end{Verbatim}
\end{tabular}
%}
\label{a:distance}
\end{table}

@@

\subsection{Use of the POR graph}
inconsistency check 
consistency checks
annotate with information extracted from biographies
what if analysis

@@

\begin{table*}[bt]                                     
\centering                                            
\caption{Results comparing between NaGEMA and iTree.}
\resizebox{2.1\columnwidth}{!}{
\begin{tabular}{p{2.1cm}lp{.1cm}cccp{.1cm}cccp{.1cm}ccc}
%\multicolumn{13}{c}{NaGEMA with all refinements} \\ \cline{3-13}
& & & \multicolumn{3}{c}{segmentation} & &  \multicolumn{3}{c}{sanad boundary}  & &  \multicolumn{3}{c}{name boundary} \\
& & & recall & precision & F-score & & recall & precision & F-score & & recall & precision & F-score \\ \cline{2-14}
\multirow{3}{*}{NaGEMA }&musnad & & 1.000 & 1.000 & 1.000 & & 0.993 & 0.990 & 0.991 & & 0.999 & 0.998 & 0.999 \\
&kafi & & 1.000 & 0.920 & 0.958 & & 0.937 & 0.970 & 0.953 & & 0.993 & 0.988 & 0.990 \\
&istibsar & & 1.000 & 0.973 & 0.986 & & 0.980 & 0.950 & 0.965 & & 0.996 & 0.998 & 0.997 \\ \cline{2-14}
& & & 1.000 & 0.964 & 0.982 & & 0.970 & 0.970 & 0.970 & & 0.996 & 0.995 & 0.995 \\ \hline \hline
%& & &  &  &  & &  &  &  & &  &  & \\
%&\multicolumn{13}{c}{iTree} \\ \cline{3-13}
& & & \multicolumn{3}{c}{accept as hadith} & &  \multicolumn{3}{c}{sanad boundary}  & &  \multicolumn{3}{c}{name boundary} \\
& & & recall & precision & F-score & & recall & precision & F-score & & recall & precision & F-score \\ \cline{2-14}
\multirow{3}{*}{iTree} &musnad & & 1.000 & NA & 1.000 & & 0.997 & 0.977 & 0.987 & & 0.998 & 1.000 & 0.999  \\
&kafi & & 0.043 & NA & 0.083 & & 1.000 & 0.313 & 0.476 & & 1.000 & 1.000 & 1.000 \\
&istibsar & & 0.028 & NA & 0.054 & & 0.875 & 0.583 & 0.700 & & 1.000 & 1.000 & 1.000 \\ \cline{2-14}
& & & 0.357 & NA & 0.379 & & 0.957 & 0.624 & 0.721 & & 0.999 & 1.000 & 1.000 \\ 
 \end{tabular}                 
}                             
\normalsize                   
\label{t:hadithresults} 
\end{table*}

\begin{table*}[bt]                                     
\centering                                            
\caption{Results without punctuation and without learning. }
\resizebox{2.1\columnwidth}{!}{
\begin{tabular}{p{2.1cm}lp{.1cm}cccp{.1cm}cccp{.1cm}ccc}
& & & \multicolumn{3}{c}{segmentation} & &  \multicolumn{3}{c}{sanad boundary}  & &  \multicolumn{3}{c}{name boundary} \\
& & & recall & precision & F-score & & recall & precision & F-score & & recall & precision & F-score \\ \cline{2-14}
\multirow{3}{*}{no punctuation} 
& musnad  & & 1.000 & 1.000 & 1.000 & & 0.982 & 0.955 & 0.968 & & 1.000 & 0.998 & 0.999 \\ 
& kafi & & 1.000 & 0.793 & 0.885 & & 0.944 & 0.967 & 0.955 & & 0.996 & 0.938 & 0.966 \\
& istibsar & & 1.000 & 1.000 & 1.000 & & 0.980 & 0.931 & 0.955 & & 0.994 & 0.985 & 0.990 \\  \cline{2-14}
& &  & 1.000 & 0.931 & 0.962 & & 0.969 & 0.951 & 0.960 & & 0.997 & 0.974 & 0.985 \\ \hline \hline
%& & & \multicolumn{3}{c}{segmentation} & &  \multicolumn{3}{c}{sanad boundary}  & &  \multicolumn{3}{c}{name boundary} \\
%& & & recall & precision & F-score & & recall & precision & F-score & & recall & precision & F-score \\ \cline{2-14}
\multirow{3}{*}{no learning}
& musnad & & 1.000 & 1.000 & 1.000 & & 0.894 & 0.985 & 0.937 & & 0.995 & 1.000 & 0.998\\
& kafi & & 1.000 & 0.920 & 0.958 & & 0.937 & 0.970 & 0.953 & & 0.993 & 0.991 & 0.992\\
& istibsar & & 1.000 & 0.973 & 0.986 & & 0.977 & 0.950 & 0.963 & & 0.996 & 0.998 & 0.997\\ \cline{2-14}
 & &  & 1.000 & 0.964 & 0.982 & & 0.936 & 0.968 & 0.951 & & 0.995 & 0.996 & 0.995 \\
 \end{tabular}                 
}                             
\normalsize                   
\label{t:nopunctlearnresults} 
\end{table*}

We ran our experiments using a dual core 2.66 Ghz 64-bit processor 
with 4GB of memory using a Linux operating system. 
We report our results in Table~\ref{t:hadithresallresults} on the 
three hadith books.
NaGEMA detected narrator chains and segmented the books 
correctly with a chain accuracy above 97\% and a narrator name
accuracy above 90\%
in all cases and across all the 1986 traditions in the three 
different books with different formats.

The row ``word count'' is an indicator of the size of the data.
The %numbers in 
rows ``names'', ``narrators'', and ``chains'' denote the total
number of the corresponding entities per book.
The row ``ignored names'' reports the number of names detected 
outside the chains and thus are irrelevant to
the application. 
%In addition, the rows ``names/narrator'' and ``narrators/chain'' 
%report the average
%number of names per narrator and 
%the average number of narrators per chain respectively.

>>>>

\begin{abstract}
\end{list}
A hadith in Islamic literature is a narration from the 
prophet Mohammad related by multiple narrators.
Establishing the authenticity of a hadith is an important task
based on which Islamic scholars issue rules that affect the life
of people around the world. 
The literature lacks an automated and exhaustive mechanism of 
authentication checks. 
This is currently manual work prone to error and almost
impossible to complete in the life time of a scholar due to the
large literature available. 
In this paper, we present a tool that uses 
an Arabic morphological analyzer 
to successfully automate the
analysis of three books of hadith selected 
arbitrarily and abstract 
each book into a graph of narrators related to each other. 
\end{abstract}

@@

In this paper, we present the hadith {\em narrator graph
extractor using morphological analysis} (NaGEMA), 
a novel technique that automates
extracting the \noArRL{sanad} from the books of traditions 
into a diagram of chains of narrators. NaGEMA segments
a book into narrations, each narration into its \noArRL{matn} 
and its \noArRL{sanad}, 
and each sanad into its separate narrators. 
This work will lead to 
automate an exhaustive \noArRL{.hady_t} authentication 
effort in the future as we plan to 
automatically analyze the biography books.

@@

{\bf Sarf.~~~}
Sarf~\footnote{Sarf (name modified for blind review) }
is 
an {\em application-specific efficient
morphological analyzer} based on
non-deterministic finite state Automata 
driven by an application-specific controller customized 
by the user. 
%Each machine in Sarf takes one letter at a time as input
%and represents a valid morphological analysis.
The application-specific controller prunes false positives
early in the run by making a decision at every input letter
to control the machines.
Sarf builds on and extends the lexicon of 
Buckwalter~\shortcite{Buckwalter:02} 
with proper and location names extracted from different online 
sources~\footnote{\href{http://alasmaa.net/}{http://alasmaa.net/ }, 
\href{http://ar.wikipedia.org/}{http://ar.wikipedia.org/}}
%as well as biblical sources~\footnote{Genesis 4:17-23; 5:1-32; 9:28-10:32; 11:10-32; 25:1-4, 12-18; 36:1-37:2; Exodus 6:14-25; Ruth 4:18-22; 1 Samuel 14:49-51; 1 Chronicles 1:1-9:44; 14:3-7; 24:1; 25:1-27:22; Nehemiah 12:8-26; Matthew 1:1-16; Luke 3:23-38}.
as well as biblical sources.
%~\footnote{Genesis 4:17-23; 5:1-32; 9:28-10:32; 11:10-32; 25:1-4, 12-18; 36:1-37:2; Exodus 6:14-25; Ruth 4:18-22; 1 Samuel 14:49-51; 1 Chronicles 1:1-9:44; 14:3-7; 24:1; 25:1-27:22; Nehemiah 12:8-26; Matthew 1:1-16; Luke 3:23-38}.

@@

In this paper we make the following contributions. 
\begin{itemize}
\item We automate the segmentation of hadith books into separate narrations and extract chains of narrators into a partial order relation graph using a knowledge based approach with morphological analysis.
\item We introduce a specialized distance metric for Arabic names used in Arabic literature. 
\item We provide a working and efficient implementation of
our work. 
\end{itemize}

The rest of this paper is organized as follows. 
We describe our approach to narrator chain extraction
in Section~\ref{sec:islamic}.
We describe how we build the partial order relation
graph in Section~\ref{sec:graph}.
We discuss related work in Section~\ref{sec:related}.
We present our results in Section~\ref{sec:results} and 
we conclude in Section~\ref{sec:future}.

@@

@@

%Several researchers~\cite{Hadithopaedia:08} attempted to automate 
%the analysis of the hadith literature. 
NaGEMA uses Sarf to automate the
analysis of three books of hadith selected 
arbitrarily~\cite{IbnHanbal,AlKulayni,AlTousi}\footnote{We obtained
  the digitized books from online sources such as 
  \href{http://www.yasoob.com/}{http://www.yasoob.com/} and 
  \href{http://www.al-eman.com/}{http://www.al-eman.com/}. }.
%  Our understanding is that those books are historical documents with 
 % public IP.}.
%\setcode{utf8}
%\begin{arabtext}
%حدثنا \emphasize{هشيم}، أخبرنا \emphasize{إسماعيل بن أبي خالد}،
% عن أبي إسحاق، عن سعيد بن جبير، قال كنا
%    مع ابن عمر رضي الله عنه حيث أفاض من
%    عرفات إلى جمع فصلى بنا المغرب ومضى
%\end{arabtext}
%\setcode{standard}
NaGEMA accepts a book as input
and segments it into a vector of hadiths. 
NaGEMA segments each hadith into its
sanad and matn parts. 
NaGEMA 
detects the chain of narrators in the sanad and 
the relation that links each narrator to his ancestor and 
predecessor in the chain. 
It also detects the full name of each narrator that is
composed of several proper names with connectors
in between.

To do that NaGEMA queries Sarf for the following morphological 
categories.
\begin{itemize}
\item $\mathit{NAME}$: a proper name of a person such as \RL{'a.hmad} .
\item $\mathit{IBN}$: a special name connector meaning the son of such as \RL{ibn}.
\item $\mathit{NMC}$: other name connectors that may be any word appearing amongst the other interesting categories. 
\item $\mathit{NISBA}$: a possessive adjective that qualifies a person such 
as \RL{alma.sriy} ``the Egyptian''. 
\item $\mathit{NRC}$: a narrator connector such as
\RL{`an} ``on behalf of'', \RL{.hada_t} ``narrated'', \RL{qaal} ``said'', 
\RL{'a_hbar} ``told'' or one of their morphological equivalents. 
\end{itemize}

@@

The process also benefits from name connector 
words and narrator connector words to enhance the
accuracy of the chain extraction and to segment the 
chain accurately into narrators.
Once we meet a name connector or a narrator 
connector word we reset the word counters that we
compare against the thresholds.

@@

We implemented our hadith chain extraction 
in a finite state machine
controller that interfaces with Sarf.
Intuitively, the controller reports a valid chain 
of narrators when a sequence of names
connected by narrator connectors appears. 
%This includes the task of finding multi-word names
%such as \RL{`bd alr.hmn} often appearing as ``run-on'' words.
The controller marks the beginning of the hadith with 
the beginning of 
the current proper name sequence,
and marks the end of the hadith with the beginning of the 
next proper name sequence. 
It detects the chain of narrators as the sequence itself.

@@

The controller moves to state $\mathit{NAME\_S}$ on
$\mathit{NAME}$.
It moves to state $\mathit{NRC\_S}$ 
when Sarf reports an $\mathit{NRC}$.
State $\mathit{NMC\_S}$
indicates that the controller expects a name to appear within 
a tolerance threshold expressed by 
$\theta_{\mathit{nmc}}$ and $\lambda_{\mathit{bmc}}$.
It returns to $\mathit{NAME\_S}$ when a name is met, 
loops when still within the threshold, and 
resets to $\mathit{TEXT\_S}$ when the threshold is exceeded 
signaling that the last collection of names met do not qualify
as a narrator or a chain of narrators.

@@

\subsection{Optimizations}

In addition to the controller, 
NaGEMA uses punctuation marks to refine its analysis and 
increase the accuracy of name detection, narration detection and narrator boundaries. 
It also learns narrator names from simple patterns that 
happen within narrator connectors and that are not 
reported by Sarf as names. 
NaGEMA learns names of narrators that consist of words missing from 
the Sarf lexicon.
NaGEMA learns the narrator names based on their context as they occurred 
between narrator connectors or contained name connectors.

\section{The narrator graph}
\label{sec:graph}

@@

The directed acyclic graph (DAG) 
in Figure~\ref{f:narrators} shows a partial order 
relation (POR) between narrators.
We automatically extracted the POR from the three books of 
hadith~\cite{IbnHanbal,AlKulayni,AlTousi}.
The nodes in boxes are the \RL{matn} of the hadith, 
and the other nodes are the narrators.
Given a set of chains of narrators similar to 
Figure~\ref{f:exhadith} we formed the DAG using graph algorithms 
that merged equal names in the chains.

\subsection{Motivation}

The hadith POR is instrumental to automate
localization of narrators in biography documents and
to segment biography documents in future work.
Consider a biography that discusses a narrator.
It mentions his professors and students.
We plan to segment the biographies with a graph 
coloring algorithm
that parses the text looking for names. 
Once a name is found it locates it in the POR graph
and colors the POR node corresponding
to it.
The narrator predecessors in the POR are supposed to be
his professors and the narrator successors in the POR 
are supposed to be his students. 
Consequently, whenever a name moves away from the 
colored local cluster in the POR graph, 
we know that we are within a new biography.

Once the biographies are analyzed, we can annotate
the POR with qualifiers of the narrators that reflect
their authenticity. 
We can also annotate the POR with the locations and 
the time the narrators lived in. 
Then we can check for time and location overlap
to discover inconsistent narrations.
This motivates our interest to automatically generate 
the POR graph.

@@

\begin{table}[tb]
%\centering
%\resizebox{.9\columnwidth}{!}{
\begin{tabular} {p{6cm}}
\begin{Verbatim}[fontsize=\relsize{-2},
frame=topline,framesep=4mm,label=\fbox{Merge Algorithm},
commandchars=\\\{\}, codes={\catcode`$=3\catcode`_=8}]
Merge(Chain $C$[])
  Chain $C_c$[]=\{\}
  foreach Chain $c_1$ in $C$ 
    $C_c = C_c \cup c_1$
    foreach Chain $c_2$ in $C-C_c$
      foreach Narrator $n_1$ in $c_1$
        start = max( $n_1$.index - radius, 0)
        end = min($n_1$.index + radius, $n_2$.size)
        $n_2$ = $c_2$[start]
        while (true)
          if (Distance($n_1$, $n_2$) < $\theta$)
            mergeNodes($n_1$, $n_2$)
          if ($n_2$ = $c_2$[end])
            break
          $n_2$ = $c_2$.next($n_2$)        
\end{Verbatim}
\end{tabular}
%}
\label{t:merge}
\end{table}

The \CodeIn{Merge} Algorithm 
iterates over the chains of narrators looking 
for equivalent narrators and merges the equivalent
narrators.
The \CodeIn{radius} parameter refines the search 
to only consider narrators who belong to the same 
generation. 
For example, a narrator relating directly to the prophet
is definitely not equivalent to a narrator separated
from the prophet by ten or more narrators.

@@

@@

  $I_1$ = extractIbn( stems(n1.connectors) )
  $I_2$ = extractIbn( stems(n2.connectors) )
  $C_1$ = stems(n1.connectors - $I_1$)
  $C_2$ = stems(n2.connectors - $I_2$)
  if ($C_1 \cap$ Places $\not= C_2 \cap$ Places)
    dist += $\delta_3$
  else
    dist -= $\delta_3$

@@

  if (correctIbnCorrespondence($I_1$, $I_2$))
      dist -= $\delta_5$
  return (MAX - dist)
\end{Verbatim}
\end{tabular}
%}
\end{table}

@@

The \CodeIn{Distance} Algorithm 
takes as input two narrators and returns 
their morphological and order distance
from each other. 
The lower the distance is, the more
equivalent the narrators are.

Each narrator has several names and several name and 
narrator connectors. 
\CodeIn{Distance} computes the names of the two narrators
and stores them in $N_1$ and $N_2$ 
and checks whether they are morphologically equivalent. 
If they are not then the distance is set to $\infty$. 
Otherwise the distance is decremented by $\delta_1$. 
Then \CodeIn{Distance} checks whether the morphologically 
equivalent names appear in the same order and 
decrements the distance if so.

\CodeIn{Distance}  pays attention to the $\mathit{NISBA}$ 
words that relate to places and checks on that to make sure
that the two narrators considered equivalent are actually 
from the same place. 
It also checks whether one place such as Baghdad 
is geographically 
contained in the other such as Iraq 
(we skip that in the Algorithm for clarity). 
\CodeIn{Distance} also checks for the equivalency of the
name connectors and the correspondence of the \RL{ibn}
connectors.

The distance metric of NaGEMA performs better than       
existing metrics~\cite{Azmi-2010} as it can detect that  
\RL{`bd al-lah bn `mr} and \RL{ibn `mr} are the same person
while other distance metrics based on simple
string similarity techniques can not.

\transfalse
\novocalize
\begin{figure*}[tb]
\center{
\resizebox{1.8\columnwidth}{!}
%{ \input{iTreeMohamada}}
{\includegraphics{iTreeMohamada}}
\caption{Output of iTree fails to detect \RL{m_hmdA}.}
\label{f:mohamada}
}
\end{figure*}
\transtrue

The closest work to NaGEMA is 
iTree~\cite{Azmi-2010,iTree}, a 
tool that uses a grammar based approach to solve 
the narrator extraction problem. 
The iTree tool preprocesses the text to remove 
punctuation and diacritics and to normalize white
spaces. 
Then iTree uses a similarity-based approach
to memory based learning~\cite{Azmi-2010} 
in order to perform 
shallow parsing of the preprocessed text. 
The iTree tool uses a parser based on a context
free grammar that enumerates several ways to express
the name of the prophet, several possible words that 
might precede the name of the prophet, several possible
words that might separate two narrators, and several
words that might separate parts of the name of 
the narrators.
The iTree tool fails when a morphological variation of 
one of the 
listed words and not an exact form of it
happens in the text. 
For example, Figure~\ref{f:mohamada} extracted 
using iTree fails to detect \RL{m_hmdA} as a name of the
prophet and does not stop at it when computing the chain 
of the hadith
\transfalse
\setcode{utf8}
%\begin{arabtext}
\RL{
حدثنا وكيع، حدثني سعيد بن السائب، عن داود بن أبي عاصم الثقفي، قال
سألت ابن عمر عن الصلاة، بمنى فقال هل سمعت {\bf محمدا}، صلى الله عليه وسلم قلت
نعم وآمنت فاهتديت به قال فإنه كان يصلي بمنى ركعتين.
%\end{arabtext}
}
\setcode{standard}
\transtrue
\vocalize

%In their grammar they consider a name to be any sequence 
%of 
%legal Arabic characters.

NaGEMA does not need a prepossessing step since 
it uses morphological analysis to process the text. 
While iTree bases its analysis on the patterns 
of words that surround the narrator names, 
NaGEMA bases its analysis on the fact that narrator
names are likely to happen in localities with a high
number of proper names. 
NaGEMA uses the narrator connectors as helpers and
uses morphological analysis to compare the stems of
those words and thus NaGEMA avoids the problem of 
``parser noise words'' that iTree faces.

The iTree tool extracts successfully 86.7 percent of
the narration chains of 90 selected traditions with 
34 simple cases and 56 hard cases. 
We ran NaGEMA against the five hadith that ship with
the tool and the two hadith that are listed in the iTree
paper and NaGEMA was able to extract all narration
chains with no mistakes even on the examples where iTree
reports a ``parser noise'' problem.

As we do not have access to the selected
90 traditions iTree reports on, we claim
that NaGEMA outperforms iTree as it works on
the full text of the presented hadith books
and reports a higher success rate.

Our approach to detect and relate names to verbs
is similar to a local grammar based 
approach~\cite{Traboulsi:09} that detects names by
associating them with verbs 
such as \RL{qaal} ``said'' and \RL{'a_hbar} ``told''. 
The local grammars in~\cite{Traboulsi:09} are also
built into transducers that generate the names and
the learning phase benefits from morphological analysis. 
We differ in that we do not use a local grammar, rather
our tolerance thresholds allow a flexible pattern that is
not easy to capture in a typical regular expression.

Another attempt to Arabic named entity recognition
uses an n-gram approach and maximum 
entropy~\cite{ANERSys:07}. 
The authors use their own corpora and gazetteers to boost 
and train the system and ignore POS tags. 
They achieve major improvements when they use lexical and
morphosyntactic features and a parallel Arabic and English 
corpora to bootstrap noisy features~\cite{Benajiba:2010}.

@@

\begin{table*}[bt]                                     
\centering                                            
\caption{Results comparing between NaGEMA and iTree.}
\resizebox{2.1\columnwidth}{!}{
\begin{tabular}{p{2.1cm}lp{.1cm}cccp{.1cm}cccp{.1cm}ccc}
%\multicolumn{13}{c}{NaGEMA with all refinements} \\ \cline{3-13}
& & & \multicolumn{3}{c}{segmentation} & &  \multicolumn{3}{c}{sanad boundary}  & &  \multicolumn{3}{c}{name boundary} \\
& & & recall & precision & F-score & & recall & precision & F-score & & recall & precision & F-score \\ \cline{2-14}
\multirow{3}{*}{NaGEMA }&musnad & & 0.999 & 0.999 & 0.999 & & 0.993 & 0.990 & 0.991 & & 0.999 & 0.998 & 0.999 \\
&kafi & & 0.999 & 0.920 & 0.958 & & 0.937 & 0.970 & 0.953 & & 0.993 & 0.988 & 0.990 \\
&istibsar & & 0.999 & 0.973 & 0.986 & & 0.980 & 0.950 & 0.965 & & 0.996 & 0.998 & 0.997 \\ \cline{2-14}
& & & 0.999 & 0.964 & 0.982 & & 0.970 & 0.970 & 0.970 & & 0.996 & 0.995 & 0.995 \\ \hline \hline
%& & &  &  &  & &  &  &  & &  &  & \\
%&\multicolumn{13}{c}{iTree} \\ \cline{3-13}
& & & \multicolumn{3}{c}{accept as hadith} & &  \multicolumn{3}{c}{sanad boundary}  & &  \multicolumn{3}{c}{name boundary} \\
& & & recall & precision & F-score & & recall & precision & F-score & & recall & precision & F-score \\ \cline{2-14}
\multirow{3}{*}{iTree} &musnad & & 0.999 & NA & 0.999 & & 0.997 & 0.977 & 0.987 & & 0.998 & 0.999 & 0.999  \\
&kafi & & 0.043 & NA & 0.083 & & 0.999 & 0.313 & 0.476 & & 0.999 & 0.999 & 0.999 \\
&istibsar & & 0.028 & NA & 0.054 & & 0.875 & 0.583 & 0.700 & & 0.999 & 0.999 & 0.999 \\ \cline{2-14}
& & & 0.357 & NA & 0.379 & & 0.957 & 0.624 & 0.721 & & 0.999 & 0.999 & 0.999 \\ 
 \end{tabular}                 
}                             
\normalsize                   
\label{t:hadithresults} 
\end{table*}

To evaluate our approach, we ran NaGEMA against three books of 
hadith~\cite{IbnHanbal,AlTousi,AlKulayni} and compared the
results to iTree. 
The results are detailed in 
Tables~\ref{t:hadithresults} and \ref{t:nopunctlearnresults}. 
The numbers we present in the tables are the result of a 
manual verification of the correctness of the outputs of 
10\% of the narrations in the books under consideration.

We used recall and precision as our evaluation metrics.
We report results on three levels of accuracy: segmentation, sanad boundary and
narrator boundary. 
Segmentation accuracy refers to the success of NaGEMA in
detecting in the plain text of a hadith book the locations of the different
narrations therein.
The sanad boundary accuracy measures the
correctness of individual narrators inside the sanad. 
Finally, the narrator boundary accuracy measures
the correctness of narrator full names.

Precisely defined, segmentation recall refers to the ratio of the narrations correctly detected against the
total number of narrations present. 
Segmentation precision refers to the fraction of correctly detected narration chains compared to all the extracted
narration chains. 
Similarly, sanad boundary recall measures whether we detected all narrators
in the sanad; while sanad boundary precision measures whether we did so without introducing false positives. 
The same concept applies to the narrator boundary where we count the number of words constituting 
a valid narrator.

\subsection{NaGEMA versus iTree}

iTree does not tackle the problem of hadith segmentation. 
In contrast to NaGEMA, it does not work on a book level, 
instead it takes as input an isolated hadith. 
It assumes the hadith book is already segmented.

Consequently, for iTree to extract the narrators from the sanad of a narration, 
the user has to extract and pass the hadith manually.
Still in many cases iTree fails to recognize the structure of
the hadith and does not extract any useful information from the narration. 
This happens for example when a slight variation of the iTree grammar rules 
happens first in the hadith.
In Table~\ref{t:hadithresults}, instead of the segmentation columns 
we report the percentage of times iTree succeeded to recognize the text passed
as a hadith.
This measure is equivalent to the concept of recall; nonetheless, precision
in this context is not applicable since by default the input is required 
to represent correctly extracted narrations.

In order to evaluate the approaches used by iTree and NaGEMA without bias,
we augmented the list of names used by 
iTree with those available in the lexicon of Sarf.
%'s lexicon on top of which NaGEMA operates.
